VFTS682: a confirmed dynamical ejection?

### 2018-05-09 VFTS682.tex
no idea what's going on
introduced todo, added list of VFTS stars

\section{Introduction}
\label{sec:intro}

How do stars form is one longstanding question in astrophysics
\cite{lada:03, zinnecker:07}. It is particularly difficult for massive stars, because these are intrinsically rare
\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, evolve fast, and
remain enshrouded in their parent cloud during the formation
process. Moreover, observations of young massive stars reveal a
complicated multipliticy structure which requires
explanation \citep[][]{kobulnicky:07,mason:09,sana:11,sana:12,kiminki:12,chini:12,kobulnicky:14,almeida:17,demarco:17}. Understanding massive star formation, possibly as a
function of metallicity, is a key question given the present and upcoming
transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{}\todo{ref} which
will reveal transients associated to massive stars
evolution and death.

The second data release (DR2) from the Gaia satellite
\cite[][]{gaia:16,brown:18} allows us to test
these hypothesis using one particular star, VFTS682. 
This star is a
very massive \citep[$M_\mathrm{ZAMS}\simeq150\,M_\odot$,][]{bestenlehner:11,schneider:18} WNh5 star in the 30 Doradus region
of the Large Magellanic Cloud (LMC), and it is presently observed at a
projected distance of $\sim$$29$\,pc from the nearest cluster of
massive stars R136 \citep[][]{bestenlehner:11}. Based on the extremely high mass of this star and
its present day apparent isolation, \cite{bestenlehner:11} proposed it
might be a candidate for isolated star formation, or a ``slow runaway'' ejected
from R136 in the past. This second option is also supported by the
N-body simulations of \cite{fujii:11, banerjee:12}.   Many other
very massive stars are present in the surroundings of R136, and a more
detailed analysis on the larger sample is desirable.

Massive stars can in principle be ejected from R136 as a consequence of dynamical
interactions \citep[][]{poveda:67,leonard:91, evans:10, fujii:11,
  allison:12, oh:16}, or by the disruption of a binary by the first
core-collapse supernova
\citep[][]{zwicky:57,blaauw:61,dedonder:97,eldridge:11,renzo:18}. However,
R136 has an estimated age of $\lesssim2$\,Myr \citep[][]{sabbi:12}, which is
shorter than the shortest stellar lifetime
\citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}, so one would not expect the binary
disruption scenario to be relevant for this cluster.

@@

@@

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=\textwidth]{./figures/main_plot}  
  \caption{position and projected relative velocity to
    R136. \todo{orient properly, load  picture on background, add cone of uncertainty}}
  \label{fig:main}
\end{figure*}

@@

\begin{itemize}
\item VFTS682 is a bona fide runaway with $v\sim60\,\mathrm{km\
    s^{-1}}$ thrown out from R136. Both its speed and the age of the
  cluster are consitent with a dynamical ejection.
  \item VFTS682 comes from R136 as was expected by
  \cite{bestenlehner:11, fujii:11, banerjee:12}, so it does not
  require isolated SFH to be explained
\item apparent age tension (connect to VFTS16 as well).
\item is R136 a single young cluster or a merger
\item estimate the influence of the gravitational potential of R136,
  what is its total mass and relaxation time?
\end{itemize}

>>>>

\section{Introduction}
\label{sec:intro}
\todo{go to major star formation immediately}
\todo{sentence 2:debate of isolated or not}
\todo{paragraph 2 about }
How do stars form is one longstanding question in astrophysics
\cite{lada:03, zinnecker:07}. It is particularly difficult for massive stars, because these are intrinsically rare
\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, evolve fast, and
remain enshrouded in their parent cloud during the formation
process. Moreover, observations of young massive stars reveal a
complicated multipliticy structure which requires
explanation \citep[][]{kobulnicky:07,mason:09,sana:11,sana:12,kiminki:12,chini:12,kobulnicky:14,almeida:17,demarco:17}. Understanding massive star formation, possibly as a
function of metallicity, is a key question given the present and upcoming
transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{}\todo{ref} which
will reveal transients associated to massive stars
evolution and death.

The second data release (DR2) from the Gaia satellite
\cite[][]{gaia:16,brown:18} allows us to test
these hypothesis using one particular star, VFTS682. 
This star is a
very massive \citep[$M_\mathrm{ZAMS}\simeq150\,M_\odot$,][]{bestenlehner:11,schneider:18} WNh5 star in the 30 Doradus region
of the Large Magellanic Cloud (LMC), and it is presently observed at a
projected distance of $\sim$$29$\,pc from the nearest cluster of
massive stars R136 \citep[][]{bestenlehner:11}. Based on the extremely high mass of this star and
its present day apparent isolation, \cite{bestenlehner:11} proposed it
might be a candidate for isolated star formation, or a ``slow runaway'' ejected
from R136 in the past. This second option is also supported by the
N-body simulations of \cite{fujii:11, banerjee:12}.   Many other
very massive stars are present in the surroundings of R136, and a more
detailed analysis on the larger sample is desirable.
\todo{this paper raised the debate/interest on VFTS682}

\todo{vfts682 is interesting also because}

Massive stars can in principle be ejected from R136 as a consequence of dynamical
interactions \citep[][]{poveda:67,leonard:91, evans:10, fujii:11,
  allison:12, oh:16}, or by the disruption of a binary by the first
core-collapse supernova
\citep[][]{zwicky:57,blaauw:61,dedonder:97,eldridge:11,renzo:18}. However,
R136 has an estimated age of $\lesssim2$\,Myr \citep[][]{sabbi:12}, which is
shorter than the shortest stellar lifetime
\citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}, so one would not expect the binary
disruption scenario to be relevant for this cluster. \todo{move up}

@@

\todo{number above not trustworthy}

@@

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=\textwidth]{./figures/main_plot}  
  \caption{The red cross indicates the position of VFTS682. Blue stars
    are those we use to define the generic ``surroundings'', while the
    green dots indicate the stars we use to probe the core of R136,
    magnified in the inset. The red arrow shows the direction of the
    proper motion of VFTS682 relative to R136, and the gray shade
    indicates the uncertainty in the direction. \todo{orient properly, load  picture on background, add cone
      of uncertainty, maybe show pm for all stars in inset?}}
  \label{fig:main}
\end{figure*}

@@

\todo{compare to Lennon: time of ejection is the same? what does it
  say for cluster}
\todo{say implications for center of R136 (same stars are there), how
  many ejected given this one, was a binary ejected in the other direction?}
\todo{does this imply there is a binary in the core ?}
\begin{itemize}
\item VFTS682 is a bona fide runaway with $v\sim60\,\mathrm{km\
    s^{-1}}$ thrown out from R136. Both its speed and the age of the
  cluster are consitent with a dynamical ejection.
  \item VFTS682 comes from R136 as was expected by
  \cite{bestenlehner:11, fujii:11, banerjee:12}, so it does not
  require isolated SFH to be explained
\item apparent age tension (connect to VFTS16 as well).
\item is R136 a single young cluster or a merger
\item estimate the influence of the gravitational potential of R136,
  what is its total mass and relaxation time?
\end{itemize}
